% resume.tex
% 
% (c) 2002 Matthew Boedicker <mboedick@mboedick.org> (original author) http://mboedick.org
% (c) 2003-2007 David J. Grant <davidgrant-at-gmail.com> http://www.davidgrant.ca
% 
% This work is licensed under the Creative Commons Attribution-ShareAlike 3.0 Unported License. To view a copy of this license, visit http://creativecommons.org/licenses/by-sa/3.0/ or send a letter to Creative Commons, 171 Second Street, Suite 300, San Francisco, California, 94105, USA.

\documentclass[letterpaper,11pt]{article}

% -----------------------------------------------------------
% Margin setup

\setlength{\voffset}{0.1in}
\setlength{\paperwidth}{8.5in}
\setlength{\paperheight}{11in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\textheight}{11in}
\setlength{\textheight}{9.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\textwidth}{7in}
\setlength{\topskip}{0in}
\setlength{\oddsidemargin}{-0.25in}
\setlength{\evensidemargin}{-0.25in}
% -----------------------------------------------------------
% \usepackage{fullpage}
% \usepackage{shading}
\usepackage{xcolor}
\usepackage{mdframed}
% \textheight=9.0in
\pagestyle{empty}
\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

% -----------------------------------------------------------
% Custom commands
\newcommand{\resitem}[1]{\item #1 \vspace{-2pt}}
% \newcommand{\resheading}[1]{{\large \parashade[.9]{sharpcorners}{\textbf{#1 \vphantom{p\^{E}}}}}}
\newcommand{\resheading}[1]{\begin{mdframed}[backgroundcolor=blue!20] \bf{#1}  \end{mdframed}}
\newcommand{\ressubheading}[4]{
  \begin{tabular*}{6.5in}{l@{\extracolsep{\fill}}r}
    \textbf{#1} & #2 \\
    \textit{#3} & \textit{#4} \\
  \end{tabular*}\vspace{-6pt}}
% -----------------------------------------------------------


\begin{document}

\begin{tabular*}{7in}{l@{\extracolsep{\fill}}r}
  \textbf{\Large Bachir El Khadir}  & 609-405-1273\\
  Sherrerd Hall, 1 Charlton Street &  bkhadir@princeton.edu \\
  Princeton University Princeton NJ 08544 & \\
\end{tabular*}
\\

\vspace{0.1in}

\resheading{Education}
\begin{itemize}
  \item
  \ressubheading{Princeton University}{Princeton, NJ}{Ph.D., ORFE}{September 2015 - present}
  \begin{itemize}
    \resitem{Research interests: Polynomial optimization, applications of optimization to machine learning, relaxations of NP-hard problems}
  \end{itemize}

\item
  \ressubheading{Ecole Polytechnique}{Paris}{M.A.Sc.}{September 2012 -  August 2015}
  \begin{itemize}
    \resitem{
      Applied Mathematics and Computer Science and as key subjects studied. 
      Relevant courses: Stochastic Models in Finance, Statistical Learning and Non parametric Estimation, Mathematical Aspects and Applications of Operations Research.}
  \end{itemize}

\item
  \ressubheading{Moulay Youssef High School}{Rabat, Morocco}{Preparatory Classes}{2011-2012}
  \begin{itemize}
    \resitem{Studied for intensive preparation in Mathematics and Physics for the highly competitive entrance exams to the leading French engineering schools.}
  \end{itemize}

\end{itemize}

\resheading{Work Experience}
\begin{itemize}
\item
  \ressubheading{Susquehanna International Group}{PA}{Quant. Research  - Intern}{June 2016 - August 2016}
  \begin{itemize}
    \resitem{Worked on predicting short term change of stocks volatility.}
  \end{itemize}

\item 
  \ressubheading{JPMorgan}{London}{Quant. Research}{May 2015 - Aug. 2016}
  \begin{itemize}
    \resitem{improve the performance of the pricing system by a factor of \bf{2.5}}.
  \end{itemize}

\item
  \ressubheading{Infosys}{Hyderabad, India}{Software Engineer}{June 2014 - August 2014}
  \begin{itemize}
    \resitem{Developed a web security scanner that:
      \begin{itemize}
      \item Analyzes the content (DOM elements) of a web page and monitors HTTP traffic
      \item Comes up with policies which can enhance the security of the application.
      \end{itemize}
    }
  \end{itemize}
\end{itemize}

\resheading{Skills}
\begin{itemize}
  \item Finance Related Projects:
  \begin{itemize}
    \item Design of automated trading strategies using genetic algorithms
    \item Study of optimal market making policies using Hawkes processes to simulate the limit order book.
  \end{itemize}
\end{itemize}

\resheading{Skills}

\begin{description}
\item[Object oriented programming:] Python, C++, Matlab 
\item[Statistics:] R
\item[Database management:] MySQL 
\end{description}

\resheading{Interests}

\begin{itemize}
\item Head of IT staff of X-Projets (Junior enterprise of Ecole Polytechnique).  Worked closely with clients to analyze their requirements, their IT needs and their resources. Projects include: 
\item Optimization of Odélice, a cooking App for iPhone: Content download is 250% faster.
\item Implementation of the web interface of the crowd funding platform “Galile360”.
\item Day trading on IG plateform (CFD, Forex)
\item Attended London Model United Nations conference and acted as a delegate from Turkey: Collaborated with a working group to create a comprehensive paper on Middle East crisis.
\end{itemize}

\end{document}
